---
title: "Citibike 2014 Plots"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(tigris)
library(leaflet)
library(sp)
library(ggmap)
library(maptools)
library(broom)
library(httr)
library(rgdal)
library(tidyverse)
library(lubridate)

load("trips.RData")
```

# Citibike 2014

1. Create a data frame that has the unique name, latitude, and longitude for each Citibike station that was present in the system in July 2014

```{r citibike.1}
start <- trips %>%
  group_by(name=start_station_name,
           lat=start_station_latitude,
           lng=start_station_longitude) %>%
  summarize()

end <- trips %>%
  group_by(name=start_station_name,
           lat=start_station_latitude,
           lng=start_station_longitude) %>%
  summarize()

stations <- merge(start, end)
```

2. Make a map showing the location of each Citibike station using ggmap

```{r citibike.2}
nyc_map <- get_map(location = c(lon = -74.00, lat = 40.72),
                   maptype = "terrain",
                   zoom = 13)

ggmap(nyc_map) +
  geom_point(data = stations, aes(x = lng, y = lat))

```

3. Do the same using leaflet, adding a popup that shows the name of the station when it's clicked on

```{r citibike.3}
leaflet() %>%
  addTiles() %>%
  setView(-74.00, 40.72, zoom = 15) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addMarkers(~lng, ~lat, popup = ~name, data = stations)
  
```

4. Then do a spatial join to combine this data frame with the Pediacities NYC neighborhood shapefile data

```{r citibike.4}
r <- GET('http://data.beta.nyc//dataset/0ff93d2d-90ba-457c-9f7e-39e47bf2ac5f/resource/35dd04fb-81b3-479b-a074-a27a37888ce7/download/d085e2f8d0b54d4590b1e7d1f35594c1pediacitiesnycneighborhoods.geojson')
nyc_neighborhoods <- readOGR(content(r, 'text'), 'OGRGeoJSON', verbose = F)

summary(nyc_neighborhoods)


points <- data.frame(lat = stations$lat, lng = stations$lng)
# turn stations into a spatial data frame
points_spdf <- points
coordinates(points_spdf) <- ~lng + lat
proj4string(points_spdf) <- proj4string(nyc_neighborhoods)
matches <- over(points_spdf, nyc_neighborhoods)
points <- cbind(points, matches)
```

5. Make a map showing the number of unique Citibike stations in each neighborhood

```{r citibike.5}
stations_by_neighborhood <- points %>%
  group_by(neighborhood) %>%
  summarize(num_stations = n())

map_data <- geo_join(nyc_neighborhoods,
                     stations_by_neighborhood,
                     "neighborhood",
                     "neighborhood")
```

6. First do this using ggmap where the fill color encodes the number of stations

```{r citibike.6}

color_pal <- colorNumeric(palette = "RdBu",
                          domain = range(map_data@data$num_stations, na.rm = T))

plot_data <- tidy(nyc_neighborhoods, region="neighborhood") %>%
  left_join(., stations_by_neighborhood, by = c("id"="neighborhood")) %>%
  filter(!is.na(num_stations))

ggmap(nyc_map) +
  geom_polygon(data = plot_data, aes(x=long, y=lat, group=group, fill=num_stations), alpha=0.75)

```

7. Then do the same using leaflet, adding a popup that shows the number of stations in a neighborhood when its shape is clicked on

```{r citibike.7}

points_with_num_stations <- points %>% left_join(stations_by_neighborhood)

# http://www.emoticonswallpapers.com/avatar/games/Ness-bike.gif

experimental_icon <- makeIcon(
  iconUrl = "http://www.emoticonswallpapers.com/avatar/games/Ness-bike.gif",
  iconWidth = 24, iconHeight = 24,
  iconAnchorX = 12, iconAnchorY = 12,
)

leaflet(map_data) %>%
  addTiles() %>%
  addPolygons(fillColor = ~color_pal(num_stations), popup = ~neighborhood) %>%
  addMarkers(~lng, ~lat, icon = experimental_icon, popup = ~ sprintf("<center><b>%s</b><br># Stations: %d</center>", neighborhood, num_stations), data = points_with_num_stations) %>%
  addProviderTiles("CartoDB.Positron") %>%
  setView(-74.00, 40.72, zoom = 15)

```

8. Now create a new data frame that has the total number of trips that depart from each station at each hour of the day on July 14th

```{r citibike.8}

trips_per_hour <- trips %>% filter(ymd == '2014-07-14') %>%
  mutate(hr_of_day = hour(starttime)) %>%
  group_by(station=start_station_name, hr_of_day, lat=start_station_latitude, lng=start_station_longitude) %>%
  summarize(num_trips = n()) %>%
  data.frame()

```

9. Do a spatial join to combine this data frame with the Pediacities NYC neighborhood shapefile data

```{r citibike.9}


# turn trips_per_hour into a spatial data frame
trips_per_hour_spdf <- trips_per_hour
coordinates(trips_per_hour_spdf) <- ~lng + lat
proj4string(trips_per_hour_spdf) <- proj4string(nyc_neighborhoods)
matches <- over(trips_per_hour_spdf, nyc_neighborhoods)
trips_per_hour <- cbind(trips_per_hour, matches)

```

10. Make a ggmap plot showing the number of trips that leave from each neighborhood at 9am, 1pm, 5pm, and 10pm, faceted by hour, where each facet contains a map where the fill color encodes the number of departing trips in each neighborhood

```{r citibike.10}

# hours = 9, 13, 17, 22

trips_by_neighborhood <- trips_per_hour %>% filter(hr_of_day == 9 | hr_of_day == 13 | hr_of_day == 17 | hr_of_day == 22) %>% group_by(neighborhood, hr_of_day) %>%
  summarize(num_trips = sum(num_trips))

color_pal <- colorNumeric(palette = "RdBu",
                          domain = range(trips_by_neighborhood$num_trips, na.rm = T))

plot_data <- tidy(nyc_neighborhoods, region="neighborhood") %>%
  left_join(., trips_by_neighborhood, by = c("id"="neighborhood")) %>%
  filter(!is.na(num_trips))

ggmap(nyc_map) +
  geom_polygon(data = plot_data, aes(x=long, y=lat, group=group, fill=num_trips), alpha=0.75) +
  facet_wrap(~hr_of_day)

```
