---
title: "NYC Maps Lab"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(tigris)
library(leaflet)
library(sp)
library(ggmap)
library(maptools)
library(broom)
library(httr)
library(rgdal)
library(tidyverse)
```

# Basic Maps

## ggmap

Make basic maps with `ggmap` by specifying a location (as a string or by lat/long) and zoom level. The `geocode` function is used to convert a string to lat/long behind the scenes and tiles are pulled from Google maps.


```{r ggmap}
nyc_map <- get_map(location = c(lon = -74.00, lat = 40.71),
                   maptype = "terrain",
                   zoom = 11)
ggmap(nyc_map)
```

## leaflet

Or make interactive maps using Leaflet for R in Rstudio.

```{r leaflet}
leaflet() %>%
  addTiles() %>%
  setView(-74.00, 40.71, zoom = 12) %>%
  addProviderTiles("CartoDB.Positron")
```

# Shapefiles

## Census tracts from tigris

The `tigris` package makes it easy to grab Census shapefiles at various levels, from tracts and zipcodes all the way up to national borders. First we look up the appropriate codes.

```{r census-tracts}
lookup_code("New York", "New York")

nyc_tracts <- tracts(state = '36', county = c('061', '047', '081', '005', '085'))
summary(nyc_tracts)
plot(nyc_tracts)
```

## Neighborhoods

Census tracts are great, but they might be too detailed for many purposes, where neighborhoods might be a more familiar unit of analysis. The PUMA Census neighborhoods for New York City lump together a few distinct areas (e.g., Chelsea, Flatiron, and the West Village), but Pediacities has more fine-grained neighborhood shapefiles available as a GeoJSON file.

```{r neighborhoods}
r <- GET('http://data.beta.nyc//dataset/0ff93d2d-90ba-457c-9f7e-39e47bf2ac5f/resource/35dd04fb-81b3-479b-a074-a27a37888ce7/download/d085e2f8d0b54d4590b1e7d1f35594c1pediacitiesnycneighborhoods.geojson')
nyc_neighborhoods <- readOGR(content(r,'text'), 'OGRGeoJSON', verbose = F)

summary(nyc_neighborhoods)
```

### ggplot

`ggplot` can plot shapefiles but requires them to be converted from a spatial polygon data frame format to plain old data frames first. `broom` handles this with its tidy command. (`fortify` from the `sp` package also works.)

```{r neighborhoods.ggplot}
nyc_neighborhoods_df <- tidy(nyc_neighborhoods)

ggplot() + 
  geom_polygon(data = nyc_neighborhoods_df, aes(x=long, y=lat, group=group))

ggmap(nyc_map) +
  geom_polygon(data = nyc_neighborhoods_df, aes(x=long, y=lat, group=group),
               color = "blue",
               fill = NA)
```

### leaflet

Or we can do an interactive version which shows neighborhood names when you click on a polygon.

```{r neighborhoods.leaflet}
leaflet(nyc_neighborhoods) %>%
  addTiles() %>%
  addPolygons(popup = ~neighborhood) %>%
  addProviderTiles("CartoDB.Positron")
```

## Spatial Joins

There are also convenient functions for mapping lat/long pairs into polygons. This is useful if you want to know what neighborhood a certain point is in, for instance.

To see how this works, let's first generate 10 random points in Manhattan.

```{r spatialjoins.1}
set.seed(42)
lats <- 40.7544882 + rnorm(10) / 100
lngs <- -73.9879923 + rnorm(10) / 200
points <- data.frame(lat=lats, lng=lngs)
points
```

In order to do the join, we first need to convert the points data frame to a spatial data frame. The `coordinates` function specifies which columns should be used for positioning, and the `proj4string` function specifies what type of projection should be used. In this case we just want a projection that's consistent with the neighborhood shapes, so we copy it over. Finally, we use the matches function to do the spatial join and bind the columns back together.

```{r spatialjoins.2}
points_spdf <- points
coordinates(points_spdf) <- ~lng + lat
proj4string(points_spdf) <- proj4string(nyc_neighborhoods)
matches <- over(points_spdf, nyc_neighborhoods)
points <- cbind(points, matches)
points
```

We can verify the matches by poking around on an interactive map and clicking the polygons and markers to see that proper neighborhoods have been assigned.

```{r spatialjoins.3}
leaflet(nyc_neighborhoods) %>%
  addTiles() %>%
  addPolygons(popup = ~neighborhood) %>%
  addMarkers(~lng, ~lat, popup = ~neighborhood, data = points) %>%
  addProviderTiles("CartoDB.Positron") %>%
  setView(-73.98, 40.75, zoom = 13)
```

Finally, we can use other data to color our polygons. As an example, we'll count the number of points in each polygon and map that to a color. The `geo_join` function makes it easy to join the spatial polygon data frame to a regular data frame. Here's the plot with Leaflet.

```{r spatialjoins.4}
points_by_neighborhood <- points %>%
  group_by(neighborhood) %>%
  summarize(num_points = n())

map_data <- geo_join(nyc_neighborhoods, points_by_neighborhood, "neighborhood", "neighborhood")

pal <- colorNumeric(palette = "RdBu",
                    domain = range(map_data@data$num_points, na.rm = T))

leaflet(map_data) %>%
  addTiles() %>%
  addPolygons(fillColor = ~pal(num_points), popup = ~neighborhood) %>%
  addMarkers(~lng, ~lat, popup = ~neighborhood, data = points) %>%
  addProviderTiles("CartoDB.Positron") %>%
  setView(-73.98, 40.75, zoom = 13)
```

We can do the same thing with `ggplot` and `ggmap` for a different look/feel, but we need to construct the underlying data frame in a slightly different way. First we use `broom` to convert the spatial data frame to a tidy data frame by neighborhood, then we join the number of points by neighborhood to the tidy data frame, and plot.

```{r spatialjoins.5}
plot_data <- tidy(nyc_neighborhoods, region="neighborhood") %>%
  left_join(., points_by_neighborhood, by=c("id"="neighborhood")) %>%
  filter(!is.na(num_points))
```


