---
title: "Citibike 2014 Plots"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(tidyverse)
library(lubridate)

load("trips.RData")
```

Create a function that computes historical trip times between any two stations: 

1. Take the trips dataframe and two station names as inputs

2. Return a 168-by-6 dataframe with summary statistics of trip times for each hour of the week (e.g., Monday 9am, Monday 10am, etc.), where the summary statistics include: 
  + Average number of trips in that hour
  + Average and median trip times for that hour
  + Standard deviation in trip time for that hour
  + Upper and lower quartiles of trip time for that hour
  
```{r citibike.1}
compute.historical.trip.times <- function(start, end) {
  
  data <- trips %>%
    filter(start_station_name == start, end_station_name == end) %>%
    mutate(week = week(ymd),
           day = weekdays(ymd),
           hour = hour(starttime)) %>%
    select(week, day, hour, start = start_station_name, end = end_station_name, tripduration)
  
  
  stats <- data %>% group_by(day, hour) %>%
    summarize(avg_trips = mean(n()),
              avg_trip_duration = mean(tripduration),
              median_trip_duration = median(tripduration),
              sd_trip_time = sd(tripduration),
              upper_quartile = quantile(tripduration, 0.75),
              lower_quartile = quantile(tripduration, 0.25))

  return(stats)
}
```

3. Use this function on trips between Penn Station and Grand Central (you can use the most popular station at each location)
"8 Ave & W 31 St"
"E 42 St & Vanderbilt Ave"

```{r citibike.2}

result = compute.historical.trip.times("8 Ave & W 31 St", "E 43 St & Vanderbilt Ave")

```

4. Make a plot of the results, where each facet is a day of the week, the x axis shows hour of the day, and the y axis shows average trip time, with transparent ribbons to show the standard deviation in trip time around the mean

```{r citibike.4}

ggplot(data = result, aes(x = hour, y = avg_trip_duration / 60)) +
  geom_line() +
  geom_ribbon(aes(ymin = (avg_trip_duration - sd_trip_time) / 60, ymax = (avg_trip_duration + sd_trip_time) / 60), alpha = 0.25) +
  geom_point(aes(size = avg_trips)) +
  facet_wrap(~day) +
  ylab("Average Trip Duration (mins)") +
  xlab("Hour of Day") +
  xlim(c(6, 18)) +
  ylim(c(0, 30))

```