---
title: 'OJ HW3: Interpretation and Cross Validation'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(tidyverse)
library(ggplot2)

oj <- read.csv("oj.csv")
```

## 1) OJ Dataset & Store Demographics

Let's return to the orange juice dataset and investigate how store demographics are related to demand.

```{r oj-dataset}

# Take the "fully interacted" model from HW2 `(logmove ~ log(price)*brand*feat)` and add in the store demographics as linear features (e.g. + demo1 + demo2+.)
fi_model <- lm(logmove ~ log(price) * brand * feat + AGE60 + EDUC + ETHNIC + INCOME + HHLARGE + WORKWOM + HVAL150 + SSTRDIST + SSTRVOL + CPDIST5 + CPWVOL5, oj)

# What demographics are significantly (t>2) related to demand?
# ------------------------------------------------------------------------------
# AGE60, EDUC, ETHNIC, INCOME, HHLARGE, HVAL150, SSTRDIST, SSTRVOL, CPDIST5, CPWVOL5

# How much did the adjusted R-squared improve with the addition of these variables?
# ------------------------------------------------------------------------------
# Old R-squared:  0.5352 => New R-Squared: 0.5848
# it improved by 0.5


```

## 2) Focusing on Two Variables

Let's focus on two variables `HVAL150` ("percent of HHs with homes >$150K") and one of your choosing. 

```{r two-variables}

# What are the means and percentiles of each of these variables?

# HVAL150
summary(oj$HVAL150)
#     Min.  1st Qu.   Median     Mean  3rd Qu.     Max. 
# 0.002509 0.123486 0.346154 0.343766 0.528313 0.916700 

# ETHNIC
summary(oj$ETHNIC)
#    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
# 0.02425 0.04191 0.07466 0.15556 0.18776 0.99569 

# Using your coefficient estimates from the regression in 1b:
#   If we move from the median value of HVAL150 to the 75th percentile (3rd quartile), how much does `log(quantity)` change each week on average?

hval_coef <- coef(fi_model)[["HVAL150"]]
delta <- summary(oj$HVAL150)[["3rd Qu."]] - summary(oj$HVAL150)[["Median"]]
q_change = exp(delta * hval_coef) # 1.067301

# If we move from the median value of EDUC to the 75th percentile (3rd quartile),
# how much does log(quantity) change each week on average?
# ------------------------------------------------------------------------------
educ_coef <- coef(fi_model)[["EDUC"]]
delta <- summary(oj$EDUC)[["3rd Qu."]] - summary(oj$EDUC)[["Median"]]
q_change = exp(delta * educ_coef) # 1.053987

# Base on this analysis, which is the more important predictor of demand?
# ------------------------------------------------------------------------------
# You could say that it's HVAL, but the difference is so small that it could be
# negligible.


# Now let's see if these variables impact price sensitivity. Add two interaction
# terms (with logprice) to the model to test this..  (Do this quickly.) 
fi_sensitivity_model <- lm(logmove ~ log(price) * brand * feat + AGE60 + EDUC + ETHNIC + INCOME + HHLARGE + WORKWOM + HVAL150 + SSTRDIST + SSTRVOL + CPDIST5 + CPWVOL5 +  EDUC * log(price) + HVAL150 * log(price), oj)

# What are the coefficients on the interaction terms?
price_EDUC_interaction <- coef(fi_sensitivity_model)[["log(price):EDUC"]] # -0.8354468
price_HVAL150_interaction <- coef(fi_sensitivity_model)[["log(price):HVAL150"]] # 2.554478

# Recall, positive values indicate lower price sensitivity and negative values
# indicate greater price sensitivity. Do your estimates make sense based on your
# intuition?
# ------------------------------------------------------------------------------
# Yes. It makes intuitive sense that neighborhoods with richer households (HVAL150)
# would be sensitive to a rise in prices for orange juice. It also makes sense
# that between education and HVAL150, HVAL150 has a stronger relationship.
# As for EDUC, the connection is more tenuous. you could make the case that
# educated households would be aware of price drops and make buying decisions
# based off of that.

# What are the coefficient estimates on the constants HVAL150 and your variable
# of choice? How do they compare to your regression from 1b?
# ------------------------------------------------------------------------------
interaction_coef_HVAL150 <- coef(fi_sensitivity_model)[["HVAL150"]] # -1.622375
interaction_coef_ETHNIC <- coef(fi_sensitivity_model)[["ETHNIC"]] # 0.6618413

# Similar to 2b, if we move from the median value of each variable to the
# 3rd quartile, how much does elasticity change? Based on this, which is more
# important to price sensitivity?
# ------------------------------------------------------------------------------

# Price Sensitivity of HVAL150
delta <- summary(oj$HVAL150)[["3rd Qu."]] - summary(oj$HVAL150)[["Median"]]
q_change = exp(delta * interaction_coef_HVAL150) # 0.7441369

# Price Sensitivity of ETHNIC
delta <- summary(oj$ETHNIC)[["3rd Qu."]] - summary(oj$ETHNIC)[["Median"]]
q_change = exp(delta * interaction_coef_ETHNIC) # 1.07773

# Based on this, the percentage of black or hispanic households is more important
# to price sensitivity.

```

## 3) Tuna Fish Question

Create make a new dataframe which takes the previous week's prices as a variable on the same line as the current week. This would enable you to see if there is intertemporal substitution. 

There are going to be a couple of steps. First is creating a new dataframe which is like the old one except that the week variable will change by a single week
```{r tuna-fish-a}
lagged_df <- oj
lagged_df$week <- lagged_df$week + 1
# lagged_df now has NEXT week and not the current one.  If we merge this by
# weeks now, this is last week's price (e.g., "lagged price").
myvars <- c("price", "week", "brand", "store")
lagged_df <- lagged_df[myvars]
lagged <- merge(oj, lagged_df, by=c("brand", "store", "week"))
```  

Investigate the Df2 and rename the lagged store values needed for a lagged price within the same store
```{r tuna-fish-b}
# NOTE: The number of observations decreased. Why? You've just lost (at least)
# one week's worth of data at each store

colnames(lagged)[18] = "lagged_price"
colnames(lagged)[6] = "price"
```

Now run a regression with this week's log(quantity) on current and last week's price.
```{r tuna-fish-c}
  
```

What do you notice about the previous week's elasticity?  Does this make sales more or less attractive from a profit maximization perspective?  Why?
```{r tuna-fish-d}
  
```

## 4) 5-Fold Cross Validation

In the last assignment you calculated the MSE on a test set.  Let's expand that code to include 5-fold cross validation.  

``` {r }

set.seed(1)
sample_percent <- .8
N <- length(oj$logmove) 
sample_size <- round(sample_percent*N)
lagged_subset <- sample_n(lagged, sample_size, replace = FALSE)
#NOTE: We can do the above three lines in a single line.  Can you think of how?  
test = anti_join(lagged, lagged_subset)

```
+ Create 5 partitions of the data of equal size.
``` {r }

set.seed(20)
folds <- 5

random_lagged <- oj[sample(nrow(oj)), ]
random_lagged$rand_obs <- seq(1, nrow(random_lagged))

random_lagged$partition <- random_lagged$rand_obs %% folds + 1
MSEs <- c(1:folds)

```
+ Create 5 training datasets using 80% of the data for each one.  This can be done multiple ways (e.g., "appending" the data together using `rbind`, randomly creating partitions and sub-setting data according to them, etc.)
``` {r }


for (i in 1:folds) {
  oj_test1 <- random_oj[which(random_oj$partition == i), ]
  oj_train1 <- anti_join(random_lagged, oj_test1)
  
  # add a regression model
  reg1 <- lm(logmove ~ log(price) * brand * feat + AGE60 + EDUC + ETHNIC + INCOME + HHLARGE + WORKWOM + HVAL150 + SSTRDIST + SSTRVOL + CPDIST5 + CPWVOL5, oj)
  # predict y
  oj_test1$logmove_hat <- predict(reg1, newdata=oj_test1)
  MSE <- mean((oj_test1$logmove_hat - oj_test1$logmove)^2)
  MSEs[i] <- MSE
}

```

+ Estimate a complex model using OLS which includes `price`, `featured`, `brand`, `brand*price` and `lagged_price`, all the sociodemographic variables and interactions of `EDUC` and `HHSIZE` with price on each of the training sets then the MSE on the test sets using the predict command.
  + Calculate the MSE for each run of the model by averaging across all the MSEs.

## 5) LASSO

Now take that same model from (4) and estimate it with LASSO.  Here is some relevant code:

```r
x <- model.matrix(~ log(price) + feat + brand + brand*log(price) + … +log(lagged_price) , data= my_awesome_data) 
y <- as.numeric(as.matrix(lagged$logmove)) 
set.seed(720) 
lasso_v1 <- glmnet(x, y, alpha=1)
plot(laxxo_v1)
coef(lasso_v1, s=lasso_v1$lambda.min)

# The cross validated version of the model (with some different objects) is
# this one: 
#   lasso_v1 <- cv.glmnet(x, y, alpha=1)
#   cvfit$lambda.min
#   coef(cvfit, s = "lambda.min")
```

```{r pressure}

```

# 06.21.18

## Spread Price & Brand Data

```{r spread-price-brand}
# first we're going to gather information by store, week and spread the prices
# for each brand

prices_per_store_wk <- oj %>% group_by(store, week) %>% select(store, week, brand, price) %>% spread(brand, price)

oj_with_brand_prices <- oj %>% merge(prices_per_store_wk, by=c("store", "week"))

mm <- oj_with_brand_prices %>% filter(brand == "minute.maid")
reg_mm <- glm(logmove ~ log(dominicks) + log(minute.maid) + log(tropicana), data = mm)

tr <- oj_with_brand_prices %>% filter(brand == "tropicana")
reg_tr <- glm(logmove ~ log(dominicks) + log(minute.maid) + log(tropicana), data = tr)

do <- oj_with_brand_prices %>% filter(brand == "dominicks")
reg_do <- glm(logmove ~ log(dominicks) + log(minute.maid) + log(tropicana), data = do)


```