---
title: "The Sampling Distribution Exercises"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(lubridate)
library(tidyverse)
```

## 7.4 Solved Exercises

### 7.1

The file `"pop2.csv"` contains information associated to the
blood pressure of an imaginary population of size 100,000. The file can be found
on the internet (http://pluto.huji.ac.il/~msby/StatThink/Datasets/pop2.
csv). The variables in this file are:

  + **id**: A numerical variable. A 7 digits number that serves as a unique identifier
of the subject.
  + **sex**: A factor variable. The sex of each subject. The values are either "MALE"
or "FEMALE".
  + **age**: A numerical variable. The age of each subject.
  + **bmi**: A numerical variable. The body mass index of each subject.
  + **systolic**: A numerical variable. The systolic blood pressure of each subject.
  + **diastolic**: A numerical variable. The diastolic blood pressure of each subject.
  + **group**: A factor variable. The blood pressure category of each subject. The
values are "NORMAL" both the systolic blood pressure is within its normal
range (between 90 and 139) and the diastolic blood pressure is within its
normal range (between 60 and 89). The value is "HIGH" if either measurements
of blood pressure are above their normal upper limits and it is
"LOW" if either measurements are below their normal lower limits.
Our goal in this question is to investigate the sampling distribution of the sample
average of the variable "bmi". We assume a sample of size n = 150.

  1. Compute the population average of the variable "bmi".
  2. Compute the population standard deviation of the variable "bmi".
  3. Compute the expectation of the sampling distribution for the sample average of the variable.
  4. Compute the standard deviation of the sampling distribution for the sample
average of the variable.
  5. Identify, using simulations, the central region that contains 80% of the
sampling distribution of the sample average.
  6. Identify, using the Central Limit Theorem, an approximation of the central
region that contains 80% of the sampling distribution of the sample
average.

```{r pop2}

pop <- read.csv('pop2.csv')

# 1. Compute the population average of the variable "bmi"
pop %>% summarize(pop_bmi_avg = mean(bmi))

# 2. Compute the population standard deviation of the variable "bmi".
pop %>% summarize(pop_bmi_sd = sd(bmi))

# 3. Compute the expectation of the sampling distribution for the sample average
# of the variable.
x_bar <- rep(0, 10^5)
for(i in 1:10^5) {
  x_sample <- sample(pop$bmi, size = 150)
  x_bar[i] <- mean(x_sample)
}
mean(x_bar)

# 4. Compute the standard deviation of the sampling distribution for the sample
# average of the variable.
sd(x_bar)

# 5. Identify, using simulations, the central region that contains 80% of the
# sampling distribution of the sample average.
quantile(x_bar, c(0.1, 0.9))

# 6. Identify, using the Central Limit Theorem, an approximation of the central
# region that contains 80% of the sampling distribution of the sample
# average.
qnorm(c(0.1, 0.9), mean(x_bar), sd(x_bar))

```