---
title: "Lab 5: Naive Bayes: Toy Example"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(scales)
library(ElemStatLearn) # spam dataset
library(e1071) # implementation of naive Bayes
library(ROCR) # evaluation metrics
```

## Exercises

Suppose you have a database on four customers. You know their income and whether or not they bought your product. Create a data frame with this data.

```{r}
train <- data.frame(buy = c("yes", "no", "no", "yes"),
                    income = c("high", "high", "medium", "low"))
train
```

Using Bayes rule calculate the probability that a customer will buy your product given that he or she has high income.

```{r}
# p(yes | income = high) = (p(income=high | yes) * p(yes)) / (p(income = high))
# p(income=high | yes) = 1/2
# p(yes) = 1/2
# p(income = high) = 1/2

# p(yes | income = high) = (0.5 * 0.5) / 0.5
(0.5 * 0.5) / 0.5
```

Estimate naive Bayes model using your data above. What is the prior probability of someone buying your product? What is the probability that a customer has a high income give that he or she bought your product?

```{r}
classifier <- naiveBayes(buy ~ income, train)
classifier 
```

Using the model you estimated above predict the probability of buying given that a customer has high income. Is your result the same as the one you calculated by hand in question 1?

```{r}
test <- data.frame(income = c("high"))
test$income <- factor(test$income, levels = c("high", "medium", "low"))
test

prediction <- predict(classifier, test, type="raw")
prediction
```

Suppose you have a database on four customers. You know their gender, income and whether or not they bought your product. Create a data frame with this data.

```{r}
train <- data.frame(buy = c("yes", "no", "no", "yes"),
                    income = c("high", "high", "medium", "low"),
                    gender = c("male", "female", "female", "male"))
train
```

Using Bayes rule calculate the probability that a customer will buy your product given that he has high income and male.

```{r}
# p(buy | income = high) = (p(income=high, gender = male| buy) * p(buy)) / (p(income = high))
# p(income=high, gender = male | buy) = 1/2
# p(yes) = 1/2
# p(income=high, gender = male) = 1/4

# p(yes | income = high) = (0.5 * 0.5) / 0.25
(0.5 * 0.5) / 0.25
```

Estimate naive Bayes model using your data above. What is the prior probability of someone buying your product? What is the probability that a customer has a high income given that he or she bought your product? What is the probability that a customer is male given that he bought your product?

```{r}
classifier <- naiveBayes(buy ~ income + gender, train)
classifier
```

Using the model you estimated above, predict the probability of buying given that a customer has a high income and is male. Is your result the same as the one you calculated by hand in question 1?

```{r}
test <- data.frame(income = c("high"), gender = c("male"))
test$income <- factor(test$income, levels = c("high", "medium", "low"))
test$gender <- factor(test$gender, levels = c("male", "female"))
test

prediction <- predict(classifier, test, type="raw")
prediction
```